Citation Analysis of Computer Science

### 2019-11-21 Scientometrics/comp.tex
update

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only records in DBLP with DOIs (4,291,130) were considered and only matches with a normalized rank of greater than 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. This procedure resulted in a dataset of 634,100 DBLP publications with Scopus identifiers where 49.6\% were labeled as article and 48.8\% as conference proceeding. Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations within the dataset. This dataset is referred to as \emph{comp}.\par

 \begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{ar_cp_ratio.pdf}
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 634,100 publications from DBLP were cross-matched with Scopus and then matched to the 27 major subject areas in the ASJC (Scopus) classifications. The largest number of publications are contributed by Computer Science; Mathematics; and Engineering and then by Social Sciences; Physics and Astronomy; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles or conference proceedings and ordered on the abscissa by the ratio of conference proceedings to articles within a major subject area (cp\_count/ar\_count). For this dataset, the major subject areas of Energy; Computer Science; Medicine; Mathematics; and Engineering had the highest ratio of conference proceedings to articles while Immunology and Microbiology; Chemistry; General (multidisciplinary); Pharmacology, Toxicology, and Pharmaceutics; and Agricultural and Biological Sciences had the lowest ratio. }
\label{ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments.
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus. We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. Consequently, we generated 20 clusters using Graclus, which were labelled 0-19 (Fig 2).

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{scopus_dblp_graclus2.pdf}
% figure caption is below the figure
\caption{caption will be inserted I just haven't been able to overcome the threshold to avoid boring tasks}
\label{heatmap}       % Give a unique label
\end{figure}
For the purpose of this study, we defined, the computer science literature as a working set of all articles in the DBLP bibliography published in 1996-2015, that had a digital object identifier (DOI), and could be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). We constructed article level clusters of the computer science literature for the years 1996-2015, at a sufficiently high level of aggregation to be comparable to disciplinary classification by journal. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. To place these clusters in context, we identified ASJC categories that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification. (which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas).

Unsurprisingly, the broad category Computer Science Applications was present in 17 of 20 clusters, followed by  Electrical and Electronic Engineering in 12 clusters.

At the other extreme, 16 minor subject areas were limited to a single cluster, viz.,
Atomic and Molecular Physics and Optics; 
Biochemistry; 
Chemistry (all); 
Cognitive Neuroscience; 
Computer Vision and Pattern Recognition;
Computers in Earth Sciences;
Earth and Planetary Sciences; 
Education; 
Electronic Optical and Magnetic Materials; 
Information Systems; 
Information Systems and Management;  
Management Science and Operations Research;
Molecular Biology;
Neurology; 
Physics and Astronomy (miscellaneous); 
and Radiology Nuclear Medicine and Imaging. These 16 areas can be described as peripheral, in various degrees, to computer science. The picture that emerges is that areas central to computer science are likely to be found in multiple clusters. \textcolor{blue} {Can we match these clusters and areas's to Salton's classification?}

\textcolor{blue}{we need Tandy's input on the paragraph below. Also, do we need a table summarizing the paragraph above as a supplement Fig.1? I've stared at these data for so long that I'm burned out so please could you rewrite first, second, third, and fourth points. A fifth if you like}

Examining these data from the perspective of the generated clusters (Fig. 1 abscissa), interactions with the Life Sciences were evident in cluster 17 (Biochemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, and Molecular Biology) and cluster 10 (Artificial Intelligence, Biomedical Engineering, Cognitive Neuroscience, Computer Science Applications, and Neurology). While acknowledging previously reported inaccuracies in journal-based classification \emph{vide supra}, these results strongly suggest that the actual structure of communication within the broad area of computer science is at variance with journal-based classification, also noted by Waltman and van Eck, to wit, ``...an indication that traditional disciplines such as those just mentioned only partly reflect the actual organization of today's scientific research."~\cite{waltman_new_2012}. The aphorism of Porter, Cohen, Roessner, and Perreault (2007), `The discrepancies between disciplinary roots and research realities loom large'~\cite{porter_measuring_2007} also invokes itself. Secondly, that a correlation exists between broad journal-based classification and article clusters in terms of subject matter. Thirdly, that some areas of computer science are relatively ubiquitous. Fourth that interactions with fields outside computer science tend to be restricted to a single cluster or small numbers of clusters. We also attempted to map these 20 clusters assembled through direct citation to the CCS classification by manually matching the top 25-50 most heavily cited publications in each cluster to corresponidng categories in the CCS. This was feasible in some cases, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization but less so in the cases of other clusters. Further, the top cited papers often derived from biology and we were concerned that this small sample (top cited papers) were not representative of the cluster.

@@

>>>>

\subsection{Data}
 A stable release of the DBLP computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the DBLP dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only records in DBLP with DOIs (4,291,130) were considered and only matches with a normalized rank of greater than 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. This procedure resulted in a dataset of 634,100 DBLP publications with Scopus identifiers where 49.6\% were labeled as article and 48.8\% as conference proceeding (Fig~\ref{fig:ar_cp_annotation}). Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations within the dataset. This dataset is referred to as \emph{comp}.\par

 \begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{ar_cp_ratio.pdf}
% figure caption is below the figure
\caption{Summary of DBLP data cross-matched with Scopus. 634,100 publications from DBLP were cross-matched with Scopus and then matched to the 27 major subject areas in the ASJC (Scopus) classifications. The largest number of publications are contributed by Computer Science; Mathematics; and Engineering and then by Social Sciences; Physics and Astronomy; and Biochemistry, Genetics, and Molecular Biology. Publications were further annotated with respect to being either articles or conference proceedings and ordered on the abscissa by the ratio of conference proceedings to articles within a major subject area (cp\_count/ar\_count). For this dataset, the major subject areas of Energy; Computer Science; Medicine; Mathematics; and Engineering had the highest ratio of conference proceedings to articles while Immunology and Microbiology; Chemistry; General (multidisciplinary); Pharmacology, Toxicology, and Pharmaceutics; and Agricultural and Biological Sciences had the lowest ratio. }
\label{fig:ar_cp_annotation}       % Give a unique label
\end{figure}

@@

\paragraph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}.  We used v1.2 in our experiments.
The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus. We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. Consequently, we generated 20 clusters using Graclus, which were labeled 0-19 (Fig~\ref{fig:heatmap}).

@@

\begin{figure}[ht]
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.5]{scopus_dblp_graclus2.pdf}
% figure caption is below the figure
\caption{caption will be inserted I just haven't been able to overcome the threshold to avoid boring tasks}
\label{fig:heatmap}       % Give a unique label
\end{figure}
For the purpose of this study, our working definition of the computer science literature was all publications in the DBLP bibliography published in the years 1996-2015, that had a digital object identifier (DOI), and could also be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). 
INSERT sentence about why one should cluster this dataset. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007,subelj_clustering_2016}. We constructed article level clusters of this computer science dataset at a sufficiently high level of aggregation to avoid cognitive challenge and cross-matched them to disciplinary classification by journal, the Scopus ASJC classification. We identified ASJC categories that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification, which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas (Fig~\ref{fig:heatmap}).

Unsurprisingly, the category Computer Science Applications was present in 17 of 20 clusters, followed by  Electrical and Electronic Engineering in 12 clusters. At the other extreme, 16 minor subject areas were limited to a single cluster, viz.,
Atomic and Molecular Physics and Optics; 
Biochemistry; 
Chemistry (all); 
Cognitive Neuroscience; 
Computer Vision and Pattern Recognition;
Computers in Earth Sciences;
Earth and Planetary Sciences; 
Education; 
Electronic Optical and Magnetic Materials; 
Information Systems; 
Information Systems and Management;  
Management Science and Operations Research;
Molecular Biology;
Neurology; 
Physics and Astronomy (miscellaneous); 
and Radiology Nuclear Medicine and Imaging. These 16 areas can be described as peripheral, in various degrees, to computer science. The picture that emerges is that areas central to computer science are likely to be found in multiple clusters. We did not attempt to map these clusters to Salton and Bergmark's classification.

Examining these data from the perspective of the generated clusters (Fig. 1 abscissa), interactions with the Life Sciences were evident in cluster 17 (Biochemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, and Molecular Biology) and cluster 10 (Artificial Intelligence, Biomedical Engineering, Cognitive Neuroscience, Computer Science Applications, and Neurology). While acknowledging previously reported inaccuracies in journal-based classification \emph{vide supra}, these results strongly suggest that the actual structure of communication within the broad area of computer science is at variance with journal-based classification, also noted by Waltman and van Eck, to wit, ``...an indication that traditional disciplines such as those just mentioned only partly reflect the actual organization of today's scientific research."~\cite{waltman_new_2012}. The aphorism of Porter, Cohen, Roessner, and Perreault (2007), `The discrepancies between disciplinary roots and research realities loom large'~\cite{porter_measuring_2007} also invokes itself. Secondly, that a correlation exists between broad journal-based classification and article clusters in terms of subject matter. Thirdly, that some areas of computer science are relatively ubiquitous. Fourth that interactions with fields outside computer science tend to be restricted to a single cluster or small numbers of clusters. We also attempted to map these 20 clusters assembled through direct citation to the CCS classification by manually matching the top 25-50 most heavily cited publications in each cluster to corresponding categories in the CCS. This was feasible in some cases, clusters 0 and 1 mapped reasonably well to the top level categories Hardware and Computer Systems Organization but less so in the cases of other clusters. Further, the top cited papers often derived from biology and we were concerned that this small sample (top cited papers) were not representative of the cluster.

@@

\section{Garbage}
